QueryMed: For Querying Biomedical Data on the Web\titlenote{Interested readers are advised to visit http://code.google.com/p/querymed to learn more about QueryMed.}

### 2010-11-20 docs/www2011/sigproc-sp.tex
the final changes including the one that we submitted

QueryMed is a query builder and result set visualizer for biomedical data that allows users to easily construct and run translational medicine queries across multiple data sources. The system is accessible for users who are unfamiliar with the SPARQL query language 
or the structure of the underlying ontologies. It permits users to retrieve data from an information-rich default set of resources or to draw on their specialized domain knowledge to determine the most appropriate data sources to query.  User input is provided through an intuitive interface, translated into SPARQL~\cite{SPARQL} queries, and executed at the relevant endpoints. The results are presented in an accessible interface that allows query refinement and filtering.

@@

The quantity of publicly available data in the biomedical domain has dramatically increased in recent years. With the linked open data movement, the Semantic Web community has been very proactive in converting these rich information datasources to Resource Description Framework (RDF) triplestores~\cite{LinkingData, Yip}.  However, to exploit the abundance of biomedical data on the Semantic Web, there is a need for easy-to-use systems that do not require the end user to have knowledge of the underlying  structure of the data or of the ontologies used in describing the data. These systems should support queries that run across multiple data sources.  There is also a  need for efficient hybrid interfaces that allow browsing data resources as well as performing queries~\cite{Jentzsch}.

%Her question must be broken up into discrete pieces that can be executed individually at one data source at a time.  Since the physician must search many data sources in order to find an answer to her single question, she requires a system that can automatically run queries over multiple data sources. Also, the physician may not know SPARQL syntax, the location of the SPARQL endpoints (i.e. data sources), or the structure of the ontologies used to describe the data in the endpoints. She is likely to want an intuitive way to query and to display the query results.  Developing intuitive ways to query multiple SPARQL endpoints and to display results is both an important and a challenging problem.  Our system, QueryMed, allows users with no knowledge of the SPARQL query language or the structure of the underlying ontologies to easily run queries across multiple SPARQL endpoints.

@@

The QueryMed system is designed to allow users without specialized knowledge of database query languages to browse and query the extensive biomedical data resources available in RDF triplestores. The simple user interface of QueryMed is intuitive for users with no knowledge of SPARQL and allows the user to run queries across multiple biomedical datasources.  Queries can be run against a default list of SPARQL endpoints, or against a set of user-defined endpoints. The system automatically translates the user input into a SPARQL query for each individual endpoint, executes the query, combines the results, and returns them to the user.  The user can choose to refine the query by iteratively modifying the original search terms and by filtering the result set.  The system's advanced query functionality enables the user to construct complex logical SPARQL queries that take advantage of the underlying structure of the data.  A general overview of QueryMed architecture is shown in Figure \ref{fig:arch_details}.

\begin{figure}
\centering
\epsfig{file=images/architecture_detail.jpg, width=3in}
\caption{QueryMed Architecture}
\label{fig:arch_details}
\end{figure}

\subsection{Sample use case}

@@

Features of the QueryMed system that distinguish it from similar systems include the ability both to perform keyword queries and to construct more advanced queries taking advantage of the structure of the data, a hybrid interface that enables the user to both query and browse data, and a graphical user interface permitting the dynamic addition of endpoints.  Furthermore, the Javascript-based user interface of the QueryMed system, implemented using the JQuery library, makes our user interface particularly attractive, easy to interact with, and capable of handling a flexible range of user input.  Another unique feature of our system is the property-based advanced query interface. This interface enables users to take advantage of the structures of the underlying ontologies used to represent the data without prior knowledge of these structures.

@@

The main contributions of QueryMed are: dynamic construction of complex SPARQL queries based on intuitive user input; dynamic addition of user-specified endpoints; and ability to run queries over multiple endpoints. Because the system is flexible and easy to use, we believe that it will be valuable to the biomedical community.  We also believe that developing systems such as such as QueryMed, which make SPARQL endpoints  easily accessible to end-users, will entice more people to expose their biomedical data as linked open data, thus promoting the growth of the Linked Open Drug Data cloud.

>>>>

QueryMed is a query builder and result set visualizer for biomedical data that allows users to easily construct and run translational medicine queries across multiple data sources. The system is accessible for users who are unfamiliar with the SPARQL query language~\cite{SPARQL}
or the structure of the underlying ontologies. It permits users to retrieve data from an information-rich default set of resources or to draw on their specialized domain knowledge to determine the most appropriate data sources to query.  User input is provided through an intuitive interface, translated into SPARQL queries, and executed at the relevant endpoints. The results are presented in an accessible interface that allows query refinement and filtering.

@@

The quantity of publicly available data in the biomedical domain has dramatically increased in recent years. With the linked open data movement, the Semantic Web community has been very proactive in converting these rich information data sources to Resource Description Framework (RDF) triplestores~\cite{LinkingData, Yip}.  However, to exploit the abundance of biomedical data on the Semantic Web, there is a need for easy-to-use systems that do not require the end user to have knowledge of the underlying  structure of the data or of the ontologies used in describing the data. These systems should support queries that run across multiple datasources.  There is also a  need for efficient hybrid interfaces that allow browsing data resources as well as performing queries~\cite{Jentzsch}.

\newpage

@@

The QueryMed system is designed to allow users without specialized knowledge of query languages to browse and query the extensive biomedical data resources available in RDF triplestores. The simple user interface of QueryMed is intuitive for users with no knowledge of SPARQL and allows the user to run queries across multiple biomedical datasources.  Queries can be run against a default list of SPARQL endpoints, or against a set of user-defined endpoints. The system automatically translates the user input into a SPARQL query for each individual endpoint, executes the query, combines the results, and returns them to the user.  The user can choose to refine the query by iteratively modifying the original search terms and by filtering the result set.  The system's advanced query functionality enables the user to construct complex logical SPARQL queries that take advantage of the underlying structure of the data.  A general overview of QueryMed architecture is shown in Figure \ref{fig:arch_details}.

\begin{figure}[h]
\centering
\epsfig{file=images/architecture_detail.jpg, width=3in}
\caption{QueryMed Architecture}
\label{fig:arch_details}
\end{figure}

\newpage

\subsection{Sample Use Case}

@@

Features of the QueryMed system that distinguish it from similar systems include the ability both to perform keyword queries and to construct more advanced queries taking advantage of the structure of the data, a hybrid interface that enables the user to both query and browse data, and a graphical user interface permitting the dynamic addition of endpoints.  Furthermore, the QueryMed user interface is attractive, easy to interact with, and capable of handling a flexible range of user input.  Another unique feature of our system is the property-based advanced query interface. This interface enables users to take advantage of the structures of the underlying ontologies used to represent the data without prior knowledge of these structures.

@@

The main contributions of QueryMed are: dynamic construction of complex SPARQL queries based on intuitive user input; dynamic addition of user-specified datasources; and ability to run queries over multiple datasources. Because the system is flexible and easy to use, we believe that it will be valuable to the biomedical community.  We also believe that developing systems such as such as QueryMed, which make SPARQL endpoints  easily accessible to end-users, will entice more people to expose their biomedical data as linked open data, thus promoting the growth of the Linked Open Data cloud.
